Identification of protoneutron star g-modes in gravitational-wave data

### 2020-07-06 methods.tex
Update for section results

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/model}
 \caption{Data from 25 1D simulations {\sc AENUS-ALCAR } and {\sc CoCoNuT}  code, solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands.} \label{fig:LMVAR}
\end{figure}

To develop the method we considered the gravitational wave signal
{\tt s20-gw-10kpc} described in Section \ref{sec:simulations}, originally
sampled at 16384 Hz but resampled at 4096 Hz.
A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on
autoregressive estimates of the local spectra for successive time intervals of 
\textcolor{red}{length 200} with a \textcolor{red}{ 90\%} overlap.
The dominant emission mode corresponds to the $\mbox{}^2 g_2$-mode and we have
developed a time-frequency method to track the ridge $m(t)$ in the spectrogram,
taking into account that it is monotonically increasing as time goes.
Starting from either the left- or right-most column of the time-frequency matrix
we identify and trace the sequence of amplitude peaks within a certain frequency
band given the monotonicity constraint. Appendix \ref{app:gmode} is providing more
details on the reconstruction of the $g$ mode ridge.

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/spectrogram}
 \caption{Spectrogram of the gravitational wave signal {\tt s20-gw-10kpc} sampled at \unit[4096]{Hz}.
   The spectrogram is obtained using data streach of 200 samples overlapping at 90\%
   with each other.} \label{fig:spectrogram}
\end{figure}

We collect the instantaneous frequency $f(t_i)$ corresponding to the ridge $m(t_i)$ for
the midpoint $t_i$ of each local time interval of the spectrogram and interpolating $f(t)$
for values in between the $t_i$. We then use equation (\ref{eq:universal}) to obtain
estimates of the time evolution of the ratio together with 95\% confidence intervals.
An exemple is given in Figure \ref{fig:ratio} where the red points are the point estimates and
the grey bands represent 95\% confidence bands. The black points are the true ratio values
computed using the mass and radius values obtained from the simulation code.

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/ratio}
 \caption{} \label{fig:ratio}
\end{figure}

>>>>

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth]{plots/model}
 \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from 25 1D simulations {\sc AENUS-ALCAR} (red) and {\sc CoCoNuT} (green) code. The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering the {\sc AENUS-ALCAR} data points.} \label{fig:LMVAR}
\end{figure}

To develop the method we considered the gravitational wave signal
{\tt s20-gw-10kpc} described in Section \ref{sec:simulations}, originally
sampled at 16384 Hz but resampled at 4096 Hz.
A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on
autoregressive estimates of the local spectra for successive time intervals of 
\textcolor{red}{length 200} with a \textcolor{red}{ 90\%} overlap.
The dominant emission mode corresponds to the $\mbox{}^2 g_2$-mode. We have
developed a time-frequency method to track the ridge $m(t)$ in the spectrogram,
taking into account that it is monotonically increasing as time goes,
a property of the $\mbox{}^2 g_2$-mode.
Starting from either the left- or right-most column of the time-frequency matrix
we identify and trace the sequence of amplitude peaks within a certain frequency
band given the monotonicity constraint. Appendix \ref{app:gmode} is providing more
details on the reconstruction of the $g$ mode ridge.

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth]{plots/spectrogram}
 \caption{Spectrogram of the gravitational wave signal {\tt s20-gw-10kpc} sampled at \unit[4096]{Hz}.
   The spectrogram is obtained using data streach of 200 samples overlapping at 90\%
   with each other.} \label{fig:spectrogram}
\end{figure}

We collect the instantaneous frequency $f(t_i)$ corresponding to the ridge $m(t_i)$ for
the midpoint $t_i$ of each local time interval of the spectrogram and interpolating $f(t)$
for values in between the $t_i$. We then use equation (\ref{eq:universal}) to obtain
estimates of the time evolution of the ratio together with 95\% confidence intervals.
An exemple is given in Figure \ref{fig:ratio} where the red points are the point estimates and
the grey bands represent 95\% confidence bands. The black points are the ratio values
computed using the mass and radius values obtained from the simulation code.

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/ratio}
 \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ as function of time extracted from the $\mbox{}^2 g_2$-mode of the {\tt s20-gw-10kpc} signal (red points and the 95\% confidence belt in grey) compared to the ratio value derived from the PNS mass and radius given by the simulation code (black points).} \label{fig:ratio}
\end{figure}

### 2020-07-06 results.tex
Update for section results

\section{Detection sensitivity with Advanced gravitational wave detectors}

>>>>

\section{Detection sensitivity with Advanced gravitational wave detectors}
\label{sec:results}

To estimate how accurately we can infer the time evolution of $r=M_{\rm PNS}/R_{\rm PNS}^2$ in the
gravitational wave detector data, we have added {\tt s20-gw-10kpc} GW signal to 
Gaussian noise realisations whose power spectral density follows advanced LIGO
spectrum~\cite{aLIGOsens:2018}. 
We have varied the distances to the source, covering a large
range of distances for which a detection in second generation of gravitational wave detectors
is feasible. The source is optimally oriented with
respect to the gravitational wave detector. We are assuming a GW signal from a core collapse
phenomena has been identified in the data and that the beginning of the GW signal is known.
The data (signal embedded in noise) are whitened using the function {\it prewhiten} of the R-package TSA.
An auto-regressive model with maximal \textcolor{red}{100} coefficients has been used.

For each of the noise realisations, we reconstruct the ratio time series {$r_i$}
of length $N$ and compute two quantities that compare {$r_i$} to ratio {$r_i^0$} derived from
the PNS mass and radius generated by the simulation code that produces {\tt s20-gw-10kpc}.
measure the estimation accuracy.
The first quantity is the coverage probability ($covpbb$) which is the fraction of the
ratio {$r_i^0$} values that fall within the 95\% confidence interval of {$r_i$}.
We also compute the $precision$ value given by
\begin{equation}
precision=\sum_1^N\frac{|r_i-r_i^0|}{r_i^0}
\end{equation}

Figure \ref{fig:s20results} is showing the median of $covpbb$ and $precision$ as
function of the distance of the source as well as the confidence belt corresponding
to the median absolute deviation.

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{plots/s20-gw_10kpc16384_covpbb_prec}
 \caption{$covpbb$ and $precision$ for {\tt s20-gw-10kpc} signal embedded in aLIGO noise at different distance from the Earth. The shaded regions are given by the median absolute deviation.} \label{fig:s20results}
\end{figure}
